---
title: "week7"
author: "Ryan Wang"
date: "2024-10-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## factor levels
```{r}
x <- c("Dec", "Apr", "Jan", "Mar")
month_levels <- c(
    "Jan", "Feb", "Mar", "Apr", "May", "Jun",
    "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"
)

y <- factor(x, levels = month_levels)
y
sort(y)
```
Any values not in the level will be silently converted to NA.

## as.numeric()
```{r}
library(tidyverse)

x1_original <- c(10, 10, 10, 50, 60, 20, 20, 40)
x1_factor <- factor(x1_original)
attributes(x1_factor)

tibble(x1_original, x1_factor) %>%
    mutate(x1_numeric = as.numeric(x1_factor))
```
as.numeric() with factors: return the order number of the factor level.

```{r}
a <- factor(c("hello", "goodbye"))
a
as.numeric(a)
```
Creates factor numbers and orders them according to an alphabetical order.

## forcats
```{r}
library("forcats")
gss_cat

gss_cat %>%
    count(race)
table(gss_cat$race, useNA = "ifany")
summary(gss_cat$race)
```

```{r}
gss_cat %>%
    count(relig)
levels(gss_cat$relig)
```

```{r}
relig_summary <- gss_cat %>%
    group_by(relig) %>%
    summarise(
        tvhours = mean(tvhours, na.rm = TRUE),
        n = n()
    )

relig_summary %>%
    ggplot(aes(x = tvhours, y = relig)) +
    geom_point()
```

### fct_reorder
Reorder the levels of relig use fct_reorder(). fct_reorder(.f, .x, .fun) takes three arguments:

.f, the factor whose levels you want to modify.

.x, a numeric vector that you want to use to reorder the levels.

Optionally, .fun, a function that’s used if there are multiple values of x for each value of f. The default value is median.
```{r}
relig_summary %>%
    ggplot(aes(
        x = tvhours,
        y = fct_reorder(.f = relig, .x = tvhours)
    )) +
    geom_point()
# the same as:
relig_summary %>%
    mutate(relig = fct_reorder(relig, tvhours)) %>%
    ggplot(aes(x = tvhours, y = relig)) +
    geom_point()

rincome_summary <-
    gss_cat %>%
    group_by(rincome) %>%
    summarise(
        age = mean(age, na.rm = TRUE),
        n = n()
    )

## Original rincome order
rincome_summary %>%
    ggplot(aes(x = age, y = rincome)) +
    geom_point()

rincome_summary %>%
    ggplot(aes(age, fct_relevel(rincome, "Not applicable"))) +
    geom_point()
```

fct_relevel moves levels to the front of the line.

```{r}
library("tidyverse")
library("palmerpenguins")
bill_info <- penguins %>%
  group_by(species) %>%
  summarise(bill_length_mm = mean(bill_length_mm, na.rm = T)) %>%
  mutate(species = fct_rev(fct_reorder(.f = species, .x=bill_length_mm, .desc = T)))

bill_info %>% ggplot(aes(x=bill_length_mm, y=species)) + 
  geom_point()

```

## fct_infreq
for bar plots, you can use fct_infreq() to order levels in decreasing frequency.
```{r}
gss_cat %>%
    mutate(marital = marital %>% fct_infreq() %>% fct_rev()) %>%
    ggplot(aes(marital)) +
    geom_bar()
```

## fct_recode
the new values go on the left
the old values go on the right
```{r}
gss_cat %>%
    mutate(partyid = fct_recode(partyid,
        "Republican, strong"    = "Strong republican",
        "Republican, weak"      = "Not str republican",
        "Independent, near rep" = "Ind,near rep",
        "Independent, near dem" = "Ind,near dem",
        "Democrat, weak"        = "Not str democrat",
        "Democrat, strong"      = "Strong democrat"
    )) %>%
    count(partyid)

gss_cat %>%
    mutate(partyid = fct_recode(partyid,
        "Republican, strong"    = "Strong republican",
        "Republican, weak"      = "Not str republican",
        "Independent, near rep" = "Ind,near rep",
        "Independent, near dem" = "Ind,near dem",
        "Democrat, weak"        = "Not str democrat",
        "Democrat, strong"      = "Strong democrat",
        "Other"                 = "No answer",
        "Other"                 = "Don't know",
        "Other"                 = "Other party"
    )) %>%
    count(partyid)
```
the new values can be not unique to combine things.

## fct_collapse
```{r}
gss_cat %>%
    mutate(partyid = fct_collapse(partyid,
        "other" = c("No answer", "Don't know", "Other party"),
        "rep" = c("Strong republican", "Not str republican"),
        "ind" = c("Ind,near rep", "Independent", "Ind,near dem"),
        "dem" = c("Not str democrat", "Strong democrat")
    )) %>%
    count(partyid)
```

