---
title: "week5"
author: "Ryan Wang"
date: "2024-09-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::write_bib("rmarkdown", file = "my-refs.bib")
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
```

# Functional loops
## lapply(): 
Loop over a list and evaluate a function on each element. returns a list.
```{r}
x <- list(a = 1:5, b = rnorm(10))
x
lapply(x, mean)
```

```{r}
x <- 1:4
lapply(x, runif)
lapply(x, runif, min=0, max=10)
```
Functions that you pass to lapply() may have other arguments.

## sapply():
The sapply() function behaves similarly to lapply().
simplify the return: 
If the result is a list where every element is length 1, then a vector is returned

If the result is a list where every element is a vector of the same length (> 1), a matrix is returned.

If it can’t figure things out, a list is returned.
```{r}
x <- list(a = 1:4, b = rnorm(10), c = rnorm(20, 1), d = rnorm(100, 5))
lapply(x, mean)
sapply(x, mean)
```


## mapply(): 
multivariate version of lapply():
```{r}
x <- 1: 10
y <- 1: 10

my_sum <- function(a, b) {
    a + b
}

## Same but with an extra check to make sure that 'a' and 'b'
## have the same lengths.
my_sum <- function(a, b) {
    ## Check that a and b are of the same length
    stopifnot(length(a) == length(b))
    a + b
}

## Apply mapply() to our function my_sum() with the inputs 'x' and 'y'
mapply(my_sum, x, y)
```

# Split
```{r}
x <- c(rnorm(n=10,0), rnorm(n=10,10), rnorm(n=10,100))
f <- gl(3, 10) # generate factor levels
f
split(x, f)

lapply(split(x, f), mean)
```

## Split a dataframe
```{r}
library("datasets")
head(airquality)

s <- split(airquality, airquality$Month)
str(s)

lapply(s, function(x) {
    colMeans(x[, c("Ozone", "Solar.R", "Wind")],
             na.rm = TRUE)
})

sapply(s, function(x) {
    colMeans(x[, c("Ozone", "Solar.R", "Wind")],
             na.rm = TRUE)
})
```


